# Introduction

This document is based on a single dataset available at `./data/statscomp.csv`. 
With this dataset we ask different research questions that motivates the statistical models developed on the paper. Explanations about the models and how the data was obtained are available in the paper.

### Exploring the dataset

This dataset follows the principle of tidy data as described in https://r4ds.had.co.nz/tidy-data.html. The key idea is that every variables has its own column, and every observation has its own unique row. Throughout this document, to facilitate our modeling approach, we will modify this dataset in different ways, often resulting in non-tidy data. However every model will start from the same base tidy dataset. 

This approach will hopefully make it easier for the reader to understand from where we are starting and adopt similar strategies in their own models. Additionally, we recommend, if the reader has the opportunity to influence the data collection process, the choice of tidy data. It is often ideal for exploratory analysis, plotting, is the basis for most models, and easy to transform to be used in different models.

```{r echo=T, warning=FALSE}
d <- readr::read_csv('./data/statscomp.csv')
```

Here we are excluding a few columns to simplify our view
```{r echo=T}
kable(head(dplyr::select(d,
                         -BestArm, -Continuous, -Differentiability, -Separability, -Scalability, -Modality,-BBOB,-BaseClass, -MaxFeval, -FevalPerDimensions),
           n=10))
```


### Column definitions of the dataset

* Algorithms: *string* Algorithm used in the optimization
* CostFunction: *string* Specific cost function used. If the cost function can be instantiated in more than one dimension this name also includes the number of dimensions, e.g. SphereN10 is has the base class Sphere and the N=10 for dimensions. 
* BestArm: *string*  represents the xalgo obtained at the end of the optimization
* NumberFunctionEval: *numeric* number of times the functon was evaluated in total
* EuclideanDistance: *numeric*  ||xalgo - xoptimal||2     
* TrueRewardDifference: *numeric* falgo - foptimal
* CumulativeRegret: *numeric* total regret 
* TimeToComplete: *numeric* time taken to complete the optimization
* Continuous: *string* function properties from the Jamil and Yang survey 2013         
* Differentiability: *string* function properties from the Jamil and Yang survey 2013  
* Separability: *string* function properties from the Jamil and Yang survey 2013           
* Scalability: *string* function properties from the Jamil and Yang survey 2013       
* Modality: *string* function properties from the Jamil and Yang survey 2013
* Ndimension: *numeric* number of dimensions
* OptimizationSuccessful"
* BBOB: *boolean* is part of the BBOB 2009 functions?         
* BaseClass: *string* the benchmark function used. E.g. SphereN10 has the base class Sphere          
* SD: *numeric* gaussian noise added to the benchmark function
* MaxFeval: *numeric* maximum number of function evaluations in total
* MaxFevalPerDimensions: *numeric* maximum number of function evaluations allowed per dimensions
* FevalPerDimensions: *numeric* number of times the benchmark function was evaluated per dimensions (some algorithms might evaluate a bit less than the maximum)
* SolveAt1: *boolean* was the problem solved at precision 1  
* SolveAt1e-1"  *boolean* was the problem solved at precision 1e-1
* SolveAt1e-3"  *boolean* was the problem solved at precision 1e-3         
* SolveAt1e-6: *boolean* was the problem solved at precision 1e-6          
* SolveEarlierAt1: *numeric*    iteration number where converged to the result at precision 1
* SolveEarlierAt1e-1: *numeric* iteration number where converged to the result at precision 1e-1
* SolveEarlierAt1e-3: *numeric* iteration number where converged to the result at precision 1e-3
* SolveEarlierAt1e-6: *numeric* iteration number where converged to the result at precision 1e-6    
* simNumber: *numeric* number of the repeated measures, in the dataset, every algorithm was evaluated 10 times in each benchmark function in each condition, in this case the number goes from 0 to 9
