---
title: "Statistical Models for Benchmark Comparison"
author: "David Issa Mattos"
date: "`r format(Sys.time(), '%d %B, %Y')`"
site: bookdown::bookdown_site
documentclass: book
bibliography: "references.bib"
biblio-style: apalike
link-citations: yes
output:
  bookdown::gitbook: default

---

```{r echo=F, eval=T}
library(knitr)
library(rmdformats)
library(kableExtra)
#data processing
library(tidyverse)
library(glue)
#modeling
library(rstan)
options(mc.cores = parallel::detectCores()) #use multiple cores
rstan_options(auto_write = TRUE) # we dont need to recompile
library(coda)
#plotting
library(patchwork)
library(viridis)
library(ggthemr)#devtools::install_github('cttobin/ggthemr')
ggthemr('flat')
ggplot2::theme_replace(axis.title.y=element_blank())
ggplot2::theme_replace(axis.text.y=element_text(size=8))

knitr::opts_chunk$set(
  echo=T, 
  warning=FALSE, 
  include=T,
  cache=F,
  prompt=FALSE,
  tidy=TRUE,
  comment=NA,
  message=FALSE,
  fig.align='center')
knitr::opts_knit$set(width=75)

source("utils.R")
```


# Preface {-}

This document is an online appendix to the paper *"Statistical Models for Benchmark Comparison"* by David Issa Mattos, Jan Bosch and Helena Holmström Olsson. It shows the process to analyze the data, including data preparation, modeling and plotting for all the models described on the paper. 


## Pre-requisites {-}

We assume that the reader has some familiarity with the R environment including packages included of the `tidyverse`, such as `dplyr` and `ggplot2`. The code presented is described and fairly commented to help readers follow the modeling process. Other programming languages such as Python with `numpy`, `pandas`, `matplotlib` etc are capable of performing the same steps, but this is out of the scope of this document. For the Bayesian models, we try to minimize dependency on a specific R package such as `brms` or `rstanarm`, and therefore we discuss the model in Stan only, since it has bindings for multiple programming languages. 

## Source code {-}

The full source code is available in the repository.

* The raw data for each model (after the described data transformation) is also available for download in the `./data` folder.
* The Stan models are available in the `./stanmodels` folder.
* The `utils` folder contains some helper functions to help plotting, and generating figures etc... In the chapter *utils* we discuss each function in detail 


The environment was defined and based on the `renv` package. The `renv` package logs all the packages in the `renv.lock` file and manages installation for a specific project. For more information see !!!!!!!
To replicate this environment, after downloading this repository, type:

```{r, eval=F, echo=T}
renv::hydrate()
```

This command will download and install all the the packages use in this work. Note that it will install the packages only for this project. 


## Compiling this document {-}

This document was created with the `bookdown` package. To compile it (and run every command to generate the models, figures and etc. ) type:

```{r eval=F}
rmarkdown::render_site('./index.Rmd',  encoding = 'UTF-8');
```

## Software environment {-}

We suggest to use Stan installed in Linux. There appear to be several problems with Stan in Mac Os Catalina.
The environment used is defined in the `renv.lock` file and session information used to compile this document is 

```{r}
sessionInfo()
```

